Bumble bee competition revisited under global climate change

### 2020-05-27 analysis.Rmd
dang

# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
## group level
rare_net_gp <- function(net, n, iter) {

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>% # drawn n sites at random
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(grouplevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  }

  out <- out %>%
      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
      group_by(site, metric) %>%
      summarize(value = mean(value)) %>%
      left_join(site_data)
}

# species level
rare_net_sp <- function(net, n, iter) {

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher") %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```

@@

ggplot(dens_div, aes(elev.mean, bb.per.fl, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~phase, scales = "free") +
  ylab("BB packing") +
  xlab("Elevation")

@@

### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")
```

### group level
```{r}
#### Create rarified networks and tabulate grouplevel metrics
groupmet_rare_2012 <- rare_net_gp(net_2012, n = 9, iter = 20)
groupmet_rare_2011 <- rare_net_gp(net_2011, n = 4, iter = 20)
groupmet_rare_2010 <- rare_net_gp(net_2010, n = 3, iter = 20)

### Plot
ggplot(filter(groupmet_rare_2012, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "V.ratio.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "V.ratio.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "V.ratio.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "discrepancy.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "discrepancy.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "discrepancy.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2012, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2011, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(groupmet_rare_2010, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()
```

### species level
```{r}
### Species level
spmet_2012 <- rare_net_sp(net_2012, 9, 20) 
spmet_2012_mean <- spmet_2012 %>%
  group_by(site, elev.mean, metric) %>%
  summarize(mean.value = mean(value))

spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
  group_by(site, elev.mean, iteration) %>%
  summarize(mean.d = mean(d))
spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
  group_by(site, elev.mean, iteration) %>%
  summarize(mean.d = mean(d))

ggplot(filter(spmet_2012_mean, metric == "d"), aes(elev.mean, mean.value)) +
  geom_point() +
  geom_smooth()

ggplot(spmet_2011, aes(elev.mean, mean.d)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(spmet_2010, aes(elev.mean, mean.d)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

### network level
```{r}
#### Create rarified networks and tabulate grouplevel metrics
netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20)
netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)
netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20)

### Plot
ggplot(filter(netmet_rare_2012, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(netmet_rare_2011, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(filter(netmet_rare_2010, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth()

ggplot(netmet_rare_2012, aes(elev.mean, number.of.compartments)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2011, aes(elev.mean, number.of.compartments)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2010, aes(elev.mean, number.of.compartments)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2012, aes(elev.mean, web.asymmetry)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2011, aes(elev.mean, web.asymmetry)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)

ggplot(netmet_rare_2010, aes(elev.mean, web.asymmetry)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~iteration)
```

>>>>

# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
# species level
rare_net_sp <- function(net, n, iter) {

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      filter(freq > 1) %>% # remove singletons
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species specificity")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

@@

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      #filter(freq > 1) %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

@@

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```
### Floral turnover ~ elevation
```{r}
fl_sp_elev <- survey %>%
  group_by(site, plant.sp) %>%
  summarize(abund = sum(flower.cover)) %>%
  left_join(site_data) %>%
  semi_join(net, by = "plant.sp") %>%
  group_by(plant.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor)

fl_sp_use_elev <- net %>%
  group_by(site, elev.mean, plant.sp) %>%
  summarize(abund = n()) %>%
  group_by(plant.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor)

ggplot(fl_sp_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  #geom_smooth() +
  geom_vline(aes(xintercept = elev.floor)) +
  geom_vline(aes(xintercept = elev.ceiling)) +
  facet_wrap(~reorder(plant.sp, elev.range), scales = "free_y")

ggplot(fl_sp_use_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  #geom_smooth() +
  geom_vline(aes(xintercept = elev.floor)) +
  geom_vline(aes(xintercept = elev.ceiling)) +
  facet_wrap(~reorder(plant.sp, elev.range), scales = "free_y")
```

@@

ggplot(dens_div, aes(elev.mean, bb.per.fl)) +
  geom_point() +
  #geom_smooth(method = "lm", se = TRUE) +
  geom_smooth() +
  facet_grid(year~phase, scales = "free") +
  ylab("BB packing") +
  xlab("Elevation")

@@

### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

### Demonstrate that the distribution of interaction frequencies does not differ across elevation classes; in other words, no systematic bias is introduced to the network analysis due to uneven frequency of species with few observations
net_freq <- net %>%
  group_by(site, dayofyear, year, bb.sp, elev.mean, elev.class2) %>%
  summarize(freq = n())

ggplot(net_freq, aes(elev.class2, freq)) +
  geom_boxplot() +
  facet_wrap(~year)

### Demonstrate that, while community composition may vary across sites, there is not evidence that any species is completely excluded from any point along our elevation gradient.
bb_elev <- net %>% 
  group_by(bb.sp, elev.mean) %>%
  summarize(abund = n())

ggplot(bb_elev, aes(elev.mean, bb.sp, color = log(abund))) +
  geom_point()

ggplot(bb_elev, aes(elev.mean, abund)) +
  geom_point() +
  facet_wrap(~bb.sp, scales = "free_y")

### Species level
spmet_2012 <- rare_net_sp(net_2012, 9, 20) %>%
  mutate(year = factor(rep("2012", n())))
spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
  mutate(year = factor(rep("2011", n())))  
spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
  mutate(year = factor(rep("2010", n())))

spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)

### Network- and group-level
netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20) %>%
  mutate(year = factor(rep("2012", n())))
netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)  %>%
  mutate(year = factor(rep("2011", n())))
netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20) %>%
  mutate(year = factor(rep("2010", n())))

netmet <- bind_rows(netmet_rare_2012, 
                    netmet_rare_2011, 
                    netmet_rare_2010) %>%
  mutate(metric = factor(metric, levels = c("web.asymmetry",
                                            "weighted.NODF",
                                            "niche.overlap.HL", 
                                            "C.score.HL")))

### Plot
ggplot(spmet_2012, aes(elev.mean, value, color = metric)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_wrap(~sp, scales = "free")

ggplot(spmet_2011, aes(elev.mean, value, color = metric)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_wrap(~sp, scales = "free")

ggplot(spmet_2010, aes(elev.mean, value, color = metric)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_wrap(~sp, scales = "free")

ggplot(netmet, aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(metric~year, scales = "free")
```

### 2020-05-27 manuscript_v2.Rmd
dang

In this study, we analyze patterns of resource partitioning in a bumble- bee-wildflower network sampled along an elevation gradient in the Berchtesgadener Alps of southeastern Germany. We first analyze species-level and group-level metrics of specialization and niche overlap to determine whether the bumble bees in our study system exihibit the expected pattern of decreased specialization and increased niche overlap at high elevation. To evaluate the plausibility of intepreting specialization and niche overlap as indices of competition, we then use independent floral surveying data to determine whether observed patterns in specialization and niche overlap correspond to variation in per-capita resource availability. Finally, we use the within-year temporal resolution of our dataset to test for reciprocal population dynamics of species within sites as a direct indication of interspecific competition. From these results, we derive predictions concerning the effects of climate change on competition as a constraint on bumble bee species and communities.

@@

**Results**\
*Colony life cycle analysis.*

```{r, echo = FALSE, message = FALSE, fig.cap = "**Figure 2:** Colony life cycle phases delimited by the proportional abundances of foundress queens and new reproductives. Dashed lines indicate where foundress queen and reproductive proportional abundances cross the thresholds of 0.75 and 0.10, respectively. Binomial regression curves are fit to the proportional abundances, with a spline term added to capture the unimodal pattern of worker abundance (not used in calculating breakpoints)."}
breaks_yday_summary <- read_csv("./output/breaks_yday.csv") %>%
  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
breaks_gdd_summary <- read_csv("./output/breaks_gdd.csv") %>%
  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))

@@

@@

>>>>

In this study, we analyze patterns of resource partitioning in a bumble- bee-wildflower network sampled along an elevation gradient in the Berchtesgadener Alps of southeastern Germany. We first analyze species- and group-level network metrics of specialization and niche overlap to determine whether the bumble bees in our study system exihibit the expected pattern of decreased specialization and increased niche overlap at high elevation. To evaluate the plausibility of intepreting specialization and niche overlap as indices of competition, we then use independent floral surveying data to determine whether observed patterns in specialization and niche overlap correspond to variation in per-capita resource availability. Finally, we use the within-year temporal resolution of our dataset to test for reciprocal population dynamics of species within sites as a direct indication of interspecific competition. From these results, we derive predictions concerning the effects of climate change on competition as a constraint on bumble bee species and communities.

@@

**Results**\
*Colony life cycle analysis.*  
```{r, echo = FALSE, message = FALSE, fig.cap = "**Figure 2:** Colony life cycle phases delimited by the proportional abundances of foundress queens and new reproductives. Dashed lines indicate where foundress queen and reproductive proportional abundances cross the thresholds of 0.75 and 0.10, respectively. Binomial regression curves are fit to the proportional abundances, with a spline term added to capture the unimodal pattern of worker abundance (not used in calculating breakpoints)."}
breaks_yday_summary <- read_csv("./output/breaks_yday.csv") %>%
  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
breaks_gdd_summary <- read_csv("./output/breaks_gdd.csv") %>%
  mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))

@@

*Specialization and niche overlap*.

@@

*Reciprocal population dynamics*.
